\textbf{Advection versus diffusion in brain intraventricular transport}

### 2025-03-28 paper/main.tex
update tau definition in mms code, clean ref.bib

% I1: What is the problem - from big picture to our precise topic - why is it interesting?
Brain development and function depend on numerous regulatory
processes, including the flow of cerebrospinal fluid
(CSF)~\cite{fame2020emergence}. Produced mostly by the choroid
plexus and partly by brain interstitial fluid~\cite{Brodal2016},
CSF is a clear liquid that flows through the ventricular system and subarachnoid
space~\cite{Damkier2013}. By transporting growth factors, nutrients, neurotransmitters
and metabolites, as well as removing waste products, CSF regulates
brain development and homeostasis~\cite{Ringers2020Role, fame2020emergence, del2010ependymal}.
Disruption of the CSF circulation is detrimental to brain health and disturbs
cognitive and motor functions~\cite{Johanson2008}.
Pathological CSF flow may also lead to
disorders such as hydrocephalus, where accumulation of
CSF often requires emergency surgical
treatment~\cite{duy2022rethinking, kahle2024paediatric, wallmeier2022role},
but for which there is little agreement on optimal treatments~\cite{flannery2014pediatric}.

@@

The tangential traction $\btau = \tau\lambda(\xx) P_{\nn}(\rr)$ that represents the cilia
motion differ for the cilia regions. We heuristically determined the constant $\tau= 0.65$ mPa
and the spatially varying function $\lambda(\xx)$. The definitions of $\rr$ and $\lambda(\xx)$
for the different cilia populations are summarized in~\Cref{tab:r_vector_and_lambda_function}.
\begingroup
\setlength{\tabcolsep}{20pt}
\renewcommand{\arraystretch}{1.5}
\begin{table}[H]
    \centering
    \caption{Definitions of $\rr$ and $\lambda(\xx)$ used in the cilia tangential 
    traction boundary condition $\btau = \tau\lambda(\xx)P_{\nn}(\rr)$,
    where $\tau=0.65$ mPa. The $x-$coordinates $x_{0, i}$ and $x_{e, i}$ define 
    the cilia domain in the dorsal ($i=d$) and ventral ($i=v$) regions of 
    the middle ventricle.}\label{tab:r_vector_and_lambda_function}
    \begin{tabular}{lcc}
        \toprule
        Cilia population & $\rr$ & $\lambda(\xx)$ \\
        \midrule 
        Anterior region, anterior ventricle 
        & $-(1, 0, 1)$ & $\frac{2}{5}$ \\
        Posterior region, anterior ventricle
        & $\ \ \ (1, 0, 1)$ & 1 \\
        Dorsal region, middle ventricle     
        & $-(1, 0, 1)$
        & $\frac{11}{4}\left(\frac{x-x_{0, d}}{x_{e, d} - x_{0, d}}\right)$\\
        Ventral region, middle ventricle    
        & $\ \ \ (1, 0, 1)$
        & $\frac{1}{2}\left(1 - \frac{x-x_{0, v}}{x_{e, v} - x_{0, v}}\right)$ \\
        \bottomrule
    \end{tabular}
\end{table}
\endgroup

@@

\subsection{Numerical verification of model implementation}\label{subsec:numerical_verification}
A cylinder geometry meshed with varying degrees of refinement was used to
perform convergence analysis of the numerical methods used to solve the Stokes equations.
The cylinder had unit length and a diameter of one half length unit.
For the analysis, we used tangential traction boundary conditions on all
of the cylinder surface. Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions}
with the velocity and pressure expressions defined in~\Cref{subsec:mms}
for $\uu_{\mathrm{ex}}$ and $p_{\mathrm{ex}}$, we calculated the error of 
the velocity approximation $\uu_h$
in the $\mathbf{L}^2$ norm and $\mathbf{H}^1$ semi-norm,
and the error of the pressure approximation $p_h$ in the $L^2$ norm (\Cref{tab:cylinder_error_rates}).
Based on the error calculations,
we calculated the experimental order of convergence,
as defined in \Cref{subsec:norms_and_eoc}.
We observe convergence in all of the reported error norms. Furthermore,
we report the error in calculation of the maximum velocities, and observe
convergence in this error as well. Finally, we calculated the $L^2$ norm of
the divergence $\nabla\cdot\uu_h$ to verify the property of mass conservation.
\begin{table}[!htbp]
    \small
    \centering
    \caption{Error norms of the velocity $\uu_h$ in the
    $\mathbf{L}^2$ and $\mathbf{H}^1$ norms,
    the pressure $p_h$ in the $L^2$ norm,
    and the error in the maximum velocity magnitude $\uu_{h, {\mathrm{mag}}}$
    for mesh refinement on a cylinder mesh with the cilia-driven/no-cardiac flow model.
    Additionally, we report the $L^2$ norm of the divergence of the velocity.
    For all error norms, convergence rates calculated
    with~\Cref{eq:convergence_order_estimate} are given
    in parentheses.}\label{tab:cylinder_error_rates}
    \begin{tabular}{cc|ccccc}
        \toprule
        $h/h_0$ & \# dofs &
        $\normltwovec{\uu_h - \uu_{\mathrm{ex}}}$
        & $\Hnorm{\uu_h - \uu_{\mathrm{ex}}}$
        & $\normltwo{p_h - p_{\mathrm{ex}}}$
        & $\normlinf{u_h} -\normlinf{u_{\mathrm{ex}}}$
        & $\normltwo{\nabla\cdot\uu_h}$\\ 
        \midrule 
        1    & 8652    & \num{2.3e-02}  (--)  & \num{4.3e-01}   (--)  & \num{5.4e-01}  (--) 
        & \num{2.4e-02} (---) & \num{1.4e-07} \\

>>>>

% I1: What is the problem - from big picture to our precise topic - why is it interesting?
Brain development and function depend on numerous regulatory
processes, including the flow of cerebrospinal fluid
(CSF)~\cite{fame2020emergence}. Produced mostly by the choroid
plexus and partly by brain interstitial fluid,
CSF is a clear liquid that flows through the ventricular system and subarachnoid
space~\cite{Brodal2016, Damkier2013}. By transporting growth factors, nutrients, neurotransmitters
and metabolites, as well as removing waste products, CSF regulates
brain development and homeostasis~\cite{Ringers2020Role, fame2020emergence, del2010ependymal}.
Disruption of the CSF circulation is detrimental to brain health and disturbs
cognitive and motor functions~\cite{Johanson2008}.
Pathological CSF flow may also lead to
disorders such as hydrocephalus, where accumulation of
CSF often requires emergency surgical
treatment~\cite{duy2022rethinking, kahle2024paediatric, wallmeier2022role},
but for which there is little agreement on optimal treatments~\cite{flannery2014pediatric}.

@@

For the tangential traction $\btau = \tau\lambda(\xx) P_{\nn}(\rr)$ that represents the cilia
motion, we heuristically determined the constant $\tau= 0.65$ mPa
and the spatially varying function $\lambda(\xx)$. The definitions of $\rr$ and $\lambda(\xx)$
for the different cilia populations are summarized in~\Cref{tab:r_vector_and_lambda_function}.
\begingroup
\setlength{\tabcolsep}{20pt}
\renewcommand{\arraystretch}{1.5}
\begin{table}[H]
    \centering
    \caption{Definitions of $\rr$ and $\lambda(\xx)$ used in the cilia tangential 
    traction boundary condition $\btau = \tau\lambda(\xx)P_{\nn}(\rr)$,
    where $\tau=0.65$ mPa. The $x-$coordinates $x_{0, i}$ and $x_{e, i}$ define 
    the cilia domain in the dorsal ($i=d$) and ventral ($i=v$) regions of 
    the middle ventricle.}\label{tab:r_vector_and_lambda_function}
    \begin{tabular}{lcc}
        \toprule
        Cilia population & $\rr$ & $\lambda(\xx)$ \\
        \midrule 
        Anterior region, anterior ventricle 
        & $-(1, 0, 1)$ & $\frac{2}{5}$ \\
        Posterior region, anterior ventricle
        & $\ \ \ (1, 0, 1)$ & 1 \\
        Dorsal region, middle ventricle     
        & $-(1, 0, 1)$
        & $\frac{11}{4}\left(\frac{x-x_{0, d}}{x_{e, d} - x_{0, d}}\right)$\\
        Ventral region, middle ventricle    
        & $\ \ \ (1, 0, 1)$
        & $\frac{1}{2}\left(1 - \frac{x-x_{0, v}}{x_{e, v} - x_{0, v}}\right)$ \\
        \bottomrule
    \end{tabular}
\end{table}
\endgroup

@@

\subsection{Numerical verification of model implementation}\label{subsec:numerical_verification}
A cylinder geometry meshed with varying degrees of refinement was used to
perform convergence analysis of the numerical methods used to solve the Stokes equations.
The cylinder had unit length and a diameter of one half length unit.
For the analysis, we used tangential traction boundary conditions on all
of the cylinder surface. Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions}
with the velocity and pressure expressions defined in~\Cref{subsec:mms}
for $\uu_{\mathrm{ex}}$ and $p_{\mathrm{ex}}$, we calculated the error of 
the velocity approximation $\uu_h$
in the $\mathbf{L}^2$ norm and $\mathbf{H}^1$ semi-norm,
and the error of the pressure approximation $p_h$ in the $L^2$ norm (\Cref{tab:cylinder_error_rates}).
Based on the error calculations,
we calculated the experimental order of convergence,
as defined in \Cref{subsec:norms_and_eoc}.
We observe convergence in all of the reported error norms. Furthermore,
we report the error in calculation of the maximum velocities, and observe
convergence in this error as well. Finally, we calculated the $L^2$ norm of
the divergence $\nabla\cdot\uu_h$ to verify the property of mass conservation.
\begin{table}[!htbp]
    \small
    \centering
    \caption{Errors of the velocity approximation $\uu_h$ in the
    $\mathbf{L}^2$ and $\mathbf{H}^1$ norms,
    the pressure $p_h$ in the $L^2$ norm,
    and the error in the maximum velocity magnitude $\uu_{h, {\mathrm{mag}}}$
    for mesh refinement on a cylinder mesh with the cilia-driven/no-cardiac flow model,
    with $\mu=1$.
    Additionally, we report the $L^2$ norm of the divergence of the velocity.
    For all error norms, convergence rates calculated
    with~\Cref{eq:convergence_order_estimate} are given
    in parentheses.}\label{tab:cylinder_error_rates}
    \begin{tabular}{cc|ccccc}
        \toprule
        $h/h_0$ & \# dofs &
        $\normltwovec{\uu_h - \uu_{\mathrm{ex}}}$
        & $\Hnorm{\uu_h - \uu_{\mathrm{ex}}}$
        & $\normltwo{p_h - p_{\mathrm{ex}}}$
        & $\normlinf{u_h} -\normlinf{u_{\mathrm{ex}}}$
        & $\normltwo{\nabla\cdot\uu_h}$\\ 
        \midrule 
        1    & 8652    & \num{2.3e-02}  (--)  & \num{4.3e-01}   (--)  & \num{5.4e-01}  (--) 
        & \num{2.4e-02} (---) & \num{1.4e-07} \\
